Supplemental Materials: \\ \emph{Surface-driven registration method for the structure-informed segmentation of diffusion MR images}

### 2016-01-27 Publications/2015-NeuroImage/04-methods.tex
revised the paper

\begin{figure*}
\includestandalone[width=\linewidth]{figure02-model}
% \includegraphics[width=\linewidth]{figure04A}
% \includegraphics[width=\linewidth]{figure04B}
% \includegraphics[width=\linewidth]{figure04C}
\caption{%
Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$
  for a real dataset.
The left plot in each panel represents a kernel density estimate of the location and spread of
  each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
  the \gls*{fa} and \gls*{adc} maps.
At the top and the right margins, the marginal distributions of $\Omega_l$ on the \gls*{fa} and
  \gls*{adc} are respectively displayed.
On the right side, the sampled distribution of each region is shown in small scatter plot, in contrast to
  the complete bivariate data.
A) shows the initialization of the model, when the anatomically-correct surfaces are not
  mapped into the \gls*{dmri} space.
The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
  \gls*{adc} maps with the original (undistorted) surfaces.
Panel B shows the equivalent information, using the surfaces that are perfectly mapped into the
  distorted \gls*{dmri} space.
Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
  reduced with respect to those in A.
Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
  are minimized with respect to panel A, and the resulting segmentation is close
  to the theoretical situation presented in panel B.}\label{fig:regseg-model}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure02}
\caption{Evaluation of \regseg{} using phantom data according to the following instrumental workflow.
  1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (i.e., ``box'', ``ball'', ``L'', ``gyrus'').
  2) A ground-truth displacement field was generated as described in \autoref{sec:regseg-digital_phantoms}, and applied to warp
      $\gammaset_R$, thereby obtaining $\gammaset_\text{true}$.
  3) After being warped, $\gammaset_\text{true}$ were projected onto the corresponding discrete 3D volume and downsampled to create partial volume effects at two resolutions,
     i.e., \isores[mm]{2.0} and \isores[mm]{1.0}, thereby producing sets of tissue fractions maps.
  4) The tissue fractions were fed into a \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
     two possible resolutions.
  5) The \regseg{} tool was applied using the warped test images as multispectral moving images and $\gammaset_R$ as shape priors.
  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_\text{test}$) and $\gammaset_\text{true}$ were assessed
     visually using automatically generated visual reports and quantitatively with the Hausdorff distance between the
     corresponding surfaces.}\label{fig:regseg-evphantoms}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Experimental workflow employed to process real data from the \acrfull*{hcp}.
  1) $\gammaset_R$ were extracted from the anatomical reference (\gls*{t1} image).
  2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
    from the fieldmap with \eqref{eq:regseg-fieldmap}.
  3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
    susceptibility-derived distortions.
  Target diffusion scalars (\gls*{fa} and \gls*{adc}) were computed with the distorted data and
    stacked to feed the multivariate input required by \regseg{}.
  4) The method was run to obtain $U_\text{test} = \hat{U}_\text{true}$, i.e., the estimate of
    the ground-truth deformation.
  5) The results were evaluated visually and quantitatively.
  The arrows point to edges in the target images (light-yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
    should be aligned with a surface, showing how distortion limits the direct mapping from the structural space
    in which the contours are defined.}\label{fig:regseg-evworkflows}
\end{figure*}

>>>>

\begin{figure*}
% \includestandalone[width=\linewidth]{figure02-model}
\includegraphics[width=\linewidth]{figure02}
% \includegraphics[width=\linewidth]{figure04B}
% \includegraphics[width=\linewidth]{figure04C}
\caption{%
Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$,
  for one real dataset.
{\color{blue}%
Panel (A, left) shows the joint distribution of the \gls*{fa} and \gls*{adc} conditioned to
  the segmentation \omegaset{} defined by the surfaces $\gammaset_R$ extracted from
  the \gls*{t1} image.
The plot was generated for reference using undistorted diffusion data, and
  therefore, $\gammaset_R$ is aligned with the \gls*{fa} and the \gls*{adc}.
The problem arises when the diffusion data present deformation, and the
  contours $\gammaset_R$ do not fit within the data (A, center).
After registration with \regseg{}, the contours are mapped onto the
  diffusion data (A, right), and the joint density plot is closer to the reference
  situation.
In panel (B), the three plots in (A) are decomposed tissue-wise.
Using filled contours, the bivariate distribution of each tissue is highlighted in its designated color,
  and represented over the remaining tissues (in gray colors).
To help assessment, dashed contours in black-to-white colors represent
  the corresponding distribution in the reference plot.
The registration process optimizes the segmentation model of \regseg{},
  and thus, the distribution of each region after registration is located
  closer to that corresponding in the reference situation, the shape of the distribution
  is more similar to the reference, and their spread is also reduced.
The effects of optimization are more noticeable on the \gls*{gm} ($\Omega_\text{GM}$) and
  the \gls*{wm} ($\Omega_\text{WM}$).
Particularly, the \gls*{wm} typically shows a bimodal distribution when the contours
  \gammaset{} do not fit the data.}
% The left plot in each panel represents a kernel density estimate of the location and spread of
%   each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
%   the \gls*{fa} and \gls*{adc} maps.
% At the top and the right margins, the marginal distributions of $\Omega_l$ on the \gls*{fa} and
%   \gls*{adc} are respectively displayed.
% On the right side, the sampled distribution of each region is shown in small scatter plot, in contrast to
%   the complete bivariate data.
% A) shows the initialization of the model, when the anatomically-correct surfaces are not
%   mapped into the \gls*{dmri} space.
% The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
%   \gls*{adc} maps with the original (undistorted) surfaces.
% Panel B shows the equivalent information, using the surfaces that are perfectly mapped into the
%   distorted \gls*{dmri} space.
% Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
%   reduced with respect to those in A.
% Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
%   are minimized with respect to panel A, and the resulting segmentation is close
%   to the theoretical situation presented in panel B.}
}\label{fig:regseg-model}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Evaluation of \regseg{} using phantom data according to the following instrumental workflow.
  1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (i.e., ``box'', ``ball'', ``L'', ``gyrus'').
  2) A ground-truth displacement field was generated as described in \autoref{sec:regseg-digital_phantoms}, and applied to warp
      $\gammaset_R$, thereby obtaining $\gammaset_\text{true}$.
  3) After being warped, $\gammaset_\text{true}$ were projected onto the corresponding discrete 3D volume and downsampled to create partial volume effects at two resolutions,
     i.e., \isores[mm]{2.0} and \isores[mm]{1.0}, thereby producing sets of tissue fractions maps.
  4) The tissue fractions were fed into a \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
     two possible resolutions.
  5) The \regseg{} tool was applied using the warped test images as multispectral moving images and $\gammaset_R$ as shape priors.
  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_\text{test}$) and $\gammaset_\text{true}$ were assessed
     visually using automatically generated visual reports and quantitatively with the Hausdorff distance between the
     corresponding surfaces.}\label{fig:regseg-evphantoms}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure04}
\caption{Experimental workflow employed to process real data from the \acrfull*{hcp}.
  1) $\gammaset_R$ were extracted from the anatomical reference (\gls*{t1} image).
  2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
    from the fieldmap with \eqref{eq:regseg-fieldmap}.
  3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
    susceptibility-derived distortions.
  Target diffusion scalars (\gls*{fa} and \gls*{adc}) were computed with the distorted data and
    stacked to feed the multivariate input required by \regseg{}.
  4) The method was run to obtain $U_\text{test} = \hat{U}_\text{true}$, i.e., the estimate of
    the ground-truth deformation.
  5) The results were evaluated visually and quantitatively.
  The arrows point to edges in the target images (light-yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
    should be aligned with a surface, showing how distortion limits the direct mapping from the structural space
    in which the contours are defined.}\label{fig:regseg-evworkflows}
\end{figure*}
